### 2019-12-17 analysis/analysis_fe.Rmd
survey report

---
title: "Untitled"
author: "Stephanie"
date: "05/12/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data

```{r}
library(tidyverse)
library(readr)
fe <- read_xlsx("experiment-export.xlsx", sheet=2) %>%
  filter(!is.na(contributor)) %>%
  mutate(contributor = factor(contributor))
```

# Check IDs
```{r}
count(fe, contributor, sort=TRUE) %>% mutate(submissions = n/12) %>% filter(submissions > 1)
```

# Compare to figure eight report

```{r}
contributors <- f1463460 %>% 
  select(contributor = `_worker_id`, 
  country = `_country`,
  validation1 = please_enter_the_unique_validation_code_shown_when_you_completed_the_survey) %>% 
  mutate(contributor = as_factor(contributor))

check <- fe %>% select(contributor, consent, validation) %>% distinct(.keep_all = TRUE) %>% 
left_join(contributors)
```

```{r}
# Load Libraries
library(tidyverse)
library(lubridate)
library(broom)
library(readxl)
library(lme4)
library(ggthemes)

invthm <- theme(
    panel.background = element_rect(fill = "transparent", colour = NA), 
    plot.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.key = element_rect(fill = "transparent", colour = NA),
    text = element_text(colour = "white", size = 20),
    axis.text = element_text(colour = "white", size = 20)
  )

###############################################################################
# Check data set 
fe %>% count(contributor)

# add data replication number
replicate <- bind_cols(replicate = sort(rep(1:12, 2)), image_name = (list.files("figures/final"))) %>% 
  mutate(replicate = as.factor(replicate))

fe <-  fe %>% left_join(., replicate, by = "image_name")

fe %>% count(group, image_name)

#remove duplilcated entries due to submit button
fe <- fe %>% group_by(group, contributor, image_name) %>%
  slice(1) %>% ungroup() %>% 
  arrange(group, contributor)

image_smry <- fe %>% count(group, contributor, order, sort = TRUE)

image_smry %>% ggplot() + geom_col(aes(x = contributor, y = n, fill = order)) + coord_flip()
# This shows multiple submission of scores by some participants

# Fix for unique contributor names
fe <-  fe %>% 
  mutate(id = paste0(group, group_indices(., group, contributor)))

```